\chapter*{Acknowledgements}
\vfill
\textit{\large I dedicate this work to my family.}

\vfill

\noindent
I would like to express my deepest gratitude to my extremely patient and supportive supervisors:
\vspace{-5pt}
\begin{description}[labelsep=0.5em, align=left, labelwidth=0em, labelindent=1em, leftmargin=0em]
\item[\bf Dr Horia Stanescu]{for his continuously awe-inspiring dedication to the tireless pursuit of knowledge and beauty, and whose mentoring and understanding has meant more to me than I'd willingly admit when sober.}
\item[\bf Prof Robert Kleta]{for his inexhaustible energy and lazer-like focus that has likely scorched entire forests in order to ensure that I never get lost in the woods for too long.}
\end{description}

\vfill
\begin{spacing}{1.3}
\noindent
I would also like to give a great deal thanks to following people:

\begin{description}[labelsep=0.5em, align=left, labelwidth=0em, labelindent=1em, leftmargin=0em]
\item{Dr Monika Mozere, my good colleague and desk-mate for her fearless upbeat confidence, and my much-needed window in the biologist's world.}
\item{Dr Detlef Bockenhauer, for his expertise, kindness, and his excellent barbecues.}
\item{To the guys; Dr Enriko Klootvijk, Dr Greg Jacquillet, Dr Anselm Zdebik, Daan Viering, Chris Cheshire,and Evans Asowata.}
\item{To the girls; Vaksha Patel, Dr. Daniela Iancu, Dr. Naomi Issler, Dr. Simona Dumitriu, Anne Kesselheim, Priya Outtandy, Felice Leung, Barbera Jensen, Sanjana Gupta, and Stephanie Dufek.}
\end{description}
\end{spacing}

\vfill
\noindent
My final thanks goes to; the DeSouza's, whose various ups and downs provided a welcome distraction from my own, and whose hospitality has no bounds; and also to Aytu\u{g}, for her extremely generous kindness and support towards the end.