\pagebreak
{\noindent \bf Publications Arising From This Thesis}\\
%\cite{AJMG:AJMGA36506}
%\cite{mencacci_missense_2015}
%\cite{le_quesne_stabej_stag3_2016}
%\cite{Rowczenio05052016}

\begingroup
\renewcommand{\chapter}[2]{}%
\begin{thebibliography}{10}
\providecommand{\url}[1]{#1}
\csname url@samestyle\endcsname
\providecommand{\newblock}{\relax}
\providecommand{\bibinfo}[2]{#2}
\providecommand{\BIBentrySTDinterwordspacing}{\spaceskip=0pt\relax}
\providecommand{\BIBentryALTinterwordstretchfactor}{4}
\providecommand{\BIBentryALTinterwordspacing}{\spaceskip=\fontdimen2\font plus
\BIBentryALTinterwordstretchfactor\fontdimen3\font minus
  \fontdimen4\font\relax}
\providecommand{\BIBforeignlanguage}[2]{{%
\expandafter\ifx\csname l@#1\endcsname\relax
\typeout{** WARNING: IEEEtran.bst: No hyphenation pattern has been}%
\typeout{** loaded for the language `#1'. Using the pattern for}%
\typeout{** the default language instead.}%
\else
\language=\csname l@#1\endcsname
\fi
#2}}
\providecommand{\BIBdecl}{\relax}
%\BIBdecl

\bibitem{AJMG:AJMGA36506}
\BIBentryALTinterwordspacing
S.~Drury, C.~Boustred, \textbf{M.~Tekman}, H.~Stanescu, R.~Kleta, N.~Lench, L.~S.
  Chitty, and R.~H. Scott, ``A novel homozygous ercc5 truncating mutation in a
  family with prenatal arthrogryposis-further evidence of genotype-phenotype
  correlation,'' \emph{American Journal of Medical Genetics Part A}, vol. 164,
  no.~7, pp. 1777--1783, 2014.
\BIBentrySTDinterwordspacing

\bibitem{mencacci_missense_2015}
\BIBentryALTinterwordspacing
N.~Mencacci, I.~Rubio-Agusti, A.~Zdebik, F.~Asmus, M.~Ludtmann, M.~Ryten,
  V.~Plagnol, A.-K. Hauser, S.~a. Bandres-Ciga, C.~Bettencourt, P.~Forabosco,
  D.~Hughes, M.~Soutar, K.~Peall, H.~Morris, D.~Trabzuni, \textbf{M.~Tekman},
  H.~Stanescu, R.~Kleta, M.~Carecchio, G.~Zorzi, N.~Nardocci, B.~Garavaglia,
  E.~Lohmann, A.~Weissbach, C.~Klein, J.~Hardy, A.~Pittman, T.~Foltynie,
  A.~Abramov, T.~Gasser, K.~Bhatia, and N.~Wood, ``A {Missense} {Mutation} in
  {KCTD}17 {Causes} {Autosomal} {Dominant} {Myoclonus}-{Dystonia},''
  \emph{American Journal of Human Genetics}, vol.~96, no.~6, pp. 938--947, Jun.
  2015.
\BIBentrySTDinterwordspacing

\bibitem{le_quesne_stabej_stag3_2016}
\BIBentryALTinterwordspacing
P.~Le~Quesne~Stabej, H.~J. Williams, C.~James, \textbf{M.~Tekman}, H.~C. Stanescu,
  R.~Kleta, L.~Ocaka, F.~Lescai, H.~L. Storr, M.~Bitner~Glindzicz,
  C.~Bacchelli, G.~S. Conway, and {GOSgene}, ``{STAG}3 truncating variant as
  the cause of primary ovarian insufficiency,'' \emph{Eur J Hum Genet},
  vol.~24, no.~1, pp. 135--138, Jan. 2016.
\BIBentrySTDinterwordspacing

\bibitem{Rowczenio05052016}
\BIBentryALTinterwordspacing
D.~M. Rowczenio, D.~S. Iancu, H.~Trojer, J.~A. Gilbertson, J.~D. Gillmore,
  A.~D. Wechalekar, \textbf{M.~Tekman}, H.~C. Stanescu, R.~Kleta, T.~Lane, P.~N.
  Hawkins, and H.~J. Lachmann, ``Autosomal dominant familial mediterranean
  fever in northern european caucasians associated with deletion of p.m694
  residue-a case series and genetic exploration,'' \emph{Rheumatology}, 2016. [\textit{In Press}]
\BIBentrySTDinterwordspacing

\end{thebibliography}
\endgroup

\pagebreak
